Statistical mechanics of Wigner negativity and magic states in quantum computation

### 2021-06-08 magic_paper.tex
simplified proof of entropic bound

	We have the following family of upper bounds on the distillation rate $R := n'/n$,
	\begin{equation}
		R \leq \frac{D_{\alpha}(W_\rho||W_\tau)}{D_{\alpha}(W_{\rho'}||W_{\tau'})} = \frac{F_\alpha(W_\rho) - F_\alpha(W_\tau)}{F'_\alpha(W_{\rho'}) - F'_\alpha(W_{\tau'})},
	\end{equation}
	where $\alpha = \frac{2a}{2b-1}$ for any positive integers $a,b$ with $a \geq b$ and $\beta F_\alpha(W_\rho) := D_\alpha (W_\rho || W_\tau) - \log \Z$ is the $\alpha$--free energy. 
\end{theorem}

@@

@@

We note that we may relate our analysis to the classical single-shot entropies that remain well-defined even in the case of quasi-distributions.

@@

For $\rho = \rho_{\rm{S}}(\epsilon)$, we get
\begin{align}
	&H_{\alpha}\left(W_{\rho_{\rm{S}}(\epsilon)}\right) = \frac{1}{1-\alpha}\log \left[ 8\left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^\alpha + \left(-\frac{1}{3} + \frac{4}{9}\epsilon \right)^\alpha \right] \nonumber\\
	&= \frac{\alpha}{\alpha-1}\log 18+ \frac{1}{1-\alpha}\log \left[ 8(3-\epsilon)^\alpha + (6-8\epsilon)^\alpha \right]
\end{align}

@@

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.35]{figs/test/H_test.pdf}
    \caption{Here, the exponent is $\alpha =$ even$/$odd for $0.10 \leq \alpha \leq 0.90$. $H_\alpha$ is decreasing in $\epsilon$ (not Schur-conocave for $\alpha < 1$) and increasing in $\alpha$.
    }
    \label{fig:Htest}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.35]{figs/test/D_test.pdf}
    \caption{Here, the exponent is $\alpha =$ even$/$odd for $0.10 \leq \alpha \leq 0.90$. $D_\alpha$ is increasing in $\epsilon$ (not Schur-convex for $\alpha < 1$) and decreasing in $\alpha$ (clear due to~\cref{thm:H2D}).
    }
    \label{fig:Dtest}
\end{figure}

@@

If $\alpha =$ even$/$odd and $\alpha < 1$, then $f(w) := \frac{-w^{\alpha-1}}{\alpha-1}$ diverges at $0$ with $\lim_{w\rightarrow 0^-} = -\infty$ and $\lim_{w\rightarrow 0^+} = +\infty$.\\
Example: If $\alpha = 2/3$, then $f(-2/3) = -3.43$, $f(-1/3) = -4.33$ and $f(+1/3) = +4.33$, so $f(w)$ is nether non-increasing nor non-decreasing.
In~\cref{fig:Htest}, it looks as if it's Schur-convex though..
}

Therefore, whenever $w_i \geq w_j$, we have that 
\begin{equation}
	-w_i^{\alpha-1} \leq -w_j^{\alpha-1},
\end{equation}
which leads to
\begin{equation}
	\frac{\partial H_\alpha}{\partial w_i} \leq \frac{\partial H_\alpha}{\partial w_j}.
\end{equation}
Therefore, $H_\alpha$ is Schur-concave and
\begin{equation}
	H_\alpha(\w) \leq H_\alpha(\w'),
\end{equation}
for any quasi-distributions $\w, \w'$ that obey $\w \succ \w'$.
\end{proof}
The set $F := \{2a/(2b-1): a,b \in \mathbb{N} \text{ and } a \geq b\}$ is dense in the rationals $\mathbb{Q}_{>1}$, which in turn are dense in the reals $\mathbb{R}_{>1}$, as any rational $c/d$ with $c>d$ can be approximated by a sequence $f_n := c2^n / (d2^n-1)$ with every element $f_n$ in the set $F$.

Now we define the classical $\alpha$-Renyi divergence in terms of the Wigner quasi-distribution of any magic state $\rho$ and the Wigner probability distribution of any equilibrium state $\tau$ as
\begin{equation}
	D_\alpha(W_\rho || W_\tau) := \frac{1}{\alpha - 1} \log \sum_{\bmx} W_{\rho}(\bmx)^\alpha W_{\tau}(\bmx)^{1-\alpha},
\end{equation}
where $\alpha$ can take values of the form $2a / (2b+1)$, for non-negative integers $a,b$ to ensure that the argument of the logarithm is positive.

@@

\subsection{Renyi divergences on quasi-distributions}
We prove the following result connecting majorization to relative majorization.
\begin{lemma}
Let $\w\in \mathbb{R}^N, \w'\in \mathbb{R}^{N'}$ be quasi-distributions and $\r\in \mathbb{R}^N, \r'\in \mathbb{R}^{N'}$ probability distributions with positive rational entries given by $r_i = a_i/K$ and $r_i' = a_i'/K$ for positive integers $a_i, a_i'$ and $K = \sum_{i=1}^N a_i = \sum_{i=1}^{N'} a_i'$. 
Then,
\begin{equation}
(\w, \r) \succ (\w', \r') \mbox{ if and only if } \Gamma_{\bma} (\w) \succ \Gamma_{\bma'} (\w'),
\end{equation}
where the embedding map $\Gamma_{\bma} :\mathbb{R}^N \rightarrow \mathbb{R}^K$ is given by
\begin{equation}
	\Gamma_{\bma}(\x) := \bigoplus_{i=1}^N x_i \eta_{a_i},
\end{equation}
with $\eta_{a_i} = (1/a_i, 1/a_i, \dots, 1/a_i)$ the uniform distribution on $a_i$ elements.
\end{lemma}
\begin{proof}
	We first note that the map $\Gamma_{\bma}$ is stochastic and has a well-defined left-inverse $\Gamma^{-1}_{\bma'}: \mathbb{R}^K \rightarrow \mathbb{R}^N$ given by
\begin{equation}
	\Gamma^{-1}_{\bma}(\w) = \left( \sum_{i=1}^{a_1} w_i,  \sum_{i = a_1 +1}^{a_1 + a_2} \hspace{-5pt} w_i, \dots , \hspace{-15pt} \sum_{i=a_1 + \dots + a_{N-1}+1}^{K} \hspace{-10pt} w_i \right),
\end{equation}
that obeys $(\Gamma^{-1}_{\bma} \circ \Gamma_{\bma}) (\x) = \x$ for all $\x \in \mathbb{R}^N$.

@@

@@

	We have the following family of upper bounds on the distillation rate $R := n'/n$,
	\begin{equation}
		R \leq \frac{D_{\alpha}(W_\rho||W_\tau)}{D_{\alpha}(W_{\rho'}||W_{\tau'})} = \frac{F_\alpha(W_\rho) - F_\alpha(W_\tau)}{F'_\alpha(W_{\rho'}) - F'_\alpha(W_{\tau'})},
	\end{equation}
	where $\alpha = \frac{2a}{2b-1}$ for any positive integers $a,b$ with $a \geq b$ and $\beta F_\alpha(W_\rho) := D_\alpha (W_\rho || W_\tau) - \log \Z$ is the $\alpha$--free energy. 
\end{theorem}
\begin{proof}
	For the sake of clarity, we write $\rho_n \coloneqq \rho_S(\epsilon)^{\otimes n}$, $\rho'_m \coloneqq \rho_S(\epsilon')^{\otimes m}$, $\tau_n \coloneqq \tau^{\otimes n}$ and $\tau'_m \coloneqq \tau'^{\otimes m}$.

	We have that 
\begin{equation}	
	(W_{\rho_n}, W_{\tau_n}) \succ (W_{\rho'_m}, W_{\tau'_m})
\end{equation}	
or equivalently $\Gamma_{\bma}(W_{\rho_n}) \succ \Gamma_{\bma'}(W_{\rho'_m})$.

For the given form of $\alpha$, $H_{\alpha}$ is Schur-concave due to~\cref{thm:HSchur}, so
\begin{equation}
	H_{\alpha}(\Gamma_{\bma}(W_{\rho_n})) \leq H_{\alpha}(\Gamma_{\bma'}(W_{\rho'_m})),
\end{equation}
and we can use~\cref{thm:H2D} to rewrite the inequality as
\begin{equation}
	K - D_{\alpha}(W_{\rho_n} || W_{\tau_n}) \leq K - D_{\alpha}(W_{\rho'_m} || W_{\tau'_m}),
\end{equation}.
Due to~\cref{thm:Da_props}, we can rewrite the inequality as
\begin{equation}
	K - nD_{\alpha}(W_{\rho} || W_{\tau}) \leq K - m D_{\alpha}(W_{\rho'} || W_{\tau'}),
\end{equation}
and rearrange it to reach
\begin{equation}
	\frac{m}{n} \leq \frac{D_{\alpha}(W_\rho||W_\tau)}{D_{\alpha}(W_{\rho'}||W_{\tau'})}.
\end{equation}
\end{proof}

We now show the following simple equivalence.
\begin{theorem} A state $\rho$ has Wigner negativity if and only if $H_\alpha (W_\rho) < 0$ for some $\alpha >1$.
\end{theorem}
The proof of this is as follows.
\begin{proof} 
If we have $H_\alpha (W_\rho) <0 $ for $\alpha >1$ then this implies that $\log \sum_{\x} W_\rho (\x)^\alpha > 1$, however it is known that $H_\alpha$ is always non-negative on probability distributions and therefore $W_\rho(\x)$ must be a strict quasi-distribution with negativity.

>>>>

	We have the following family of upper bounds on the distillation rate $R := n'/n$,
	\begin{equation}
		R \leq \frac{D_{\alpha}(W_\rho||W_\tau)}{D_{\alpha}(W_{\rho'}||W_{\tau'})} = \frac{F_\alpha(W_\rho) - F_\alpha(W_\tau)}{F_\alpha(W_{\rho'}) - F_\alpha(W_{\tau'})},
	\end{equation}
	where $\alpha = \frac{2a}{2b-1}$ for any positive integers $a,b$ with $a \geq b$ and $\beta F_\alpha(W_\rho) := D_\alpha (W_\rho || W_\tau) - \log \Z$ is the $\alpha$--free energy. 
\end{theorem}

@@

We prove the following result connecting majorization to relative majorization.
\begin{lemma}
	Let $\w\in \mathbb{R}^N, \w'\in \mathbb{R}^{N'}$ be quasi-distributions and $\r\in \mathbb{R}^N, \r'\in \mathbb{R}^{N'}$ probability distributions with positive rational entries given by $r_i = a_i/K$ and $r_i' = a_i'/K$ for positive integers $a_i, a_i'$ and $K = \sum_{i=1}^N a_i = \sum_{i=1}^{N'} a_i'$. 
Then,
\begin{equation}
(\w, \r) \succ (\w', \r') \mbox{ if and only if } \Gamma_{\bma} (\w) \succ \Gamma_{\bma'} (\w'),
\end{equation}
where the embedding map $\Gamma_{\bma} :\mathbb{R}^N \rightarrow \mathbb{R}^K$ is given by
\begin{equation}
	\Gamma_{\bma}(\x) := \bigoplus_{i=1}^N x_i \eta_{a_i},
\end{equation}
with $\eta_{a_i} = (1/a_i, 1/a_i, \dots, 1/a_i)$ the uniform distribution on $a_i$ elements.
\end{lemma}
\begin{proof}
	We first note that the map $\Gamma_{\bma}$ is stochastic and has a well-defined left-inverse $\Gamma^{-1}_{\bma'}: \mathbb{R}^K \rightarrow \mathbb{R}^N$ given by
\begin{equation}
	\Gamma^{-1}_{\bma}(\w) = \left( \sum_{i=1}^{a_1} w_i,  \sum_{i = a_1 +1}^{a_1 + a_2} \hspace{-5pt} w_i, \dots , \hspace{-15pt} \sum_{i=a_1 + \dots + a_{N-1}+1}^{K} \hspace{-10pt} w_i \right),
\end{equation}
that obeys $(\Gamma^{-1}_{\bma} \circ \Gamma_{\bma}) (\x) = \x$ for all $\x \in \mathbb{R}^N$.

@@

In this section, we develop the tools that allow the use of entropic functions on quasi-probabilities to retrieve magic distillation bounds.

@@

For $\rho = \rho_{\rm{S}}(\epsilon)$, we get
\begin{align}
	&H_{\alpha}\left(W_{\rho_{\rm{S}}(\epsilon)}\right) = \frac{1}{1-\alpha}\log \left[ 8\left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^\alpha + \left(-\frac{1}{3} + \frac{4}{9}\epsilon \right)^\alpha \right] \nonumber\\
	&\ddd{= \frac{\alpha}{\alpha-1}\log 18+ \frac{1}{1-\alpha}\log \left[ 8(3-\epsilon)^\alpha + (6-8\epsilon)^\alpha \right]}
\end{align}

@@

@@

If $\alpha =$ even$/$odd and $\alpha < 1$, then $f(w) := \frac{-w^{\alpha-1}}{\alpha-1}$ diverges at $0$ with $\lim_{w\rightarrow 0^-} = -\infty$ and $\lim_{w\rightarrow 0^+} = +\infty$.\\
Example: If $\alpha = 2/3$, then $f(-2/3) = -3.43$, $f(-1/3) = -4.33$ and $f(+1/3) = +4.33$, so $f(w)$ is nether non-increasing nor non-decreasing.
}

Therefore, whenever $w_i \geq w_j$, we have that 
\begin{equation}
	-w_i^{\alpha-1} \leq -w_j^{\alpha-1},
\end{equation}
which leads to
\begin{equation}
	\frac{\partial H_\alpha}{\partial w_i} \leq \frac{\partial H_\alpha}{\partial w_j}.
\end{equation}
Therefore, $H_\alpha$ is Schur-concave and
\begin{equation}
	H_\alpha(\w) \leq H_\alpha(\w'),
\end{equation}
for any quasi-distributions $\w, \w'$ that obey $\w \succ \w'$.
\end{proof}
The set $F := \{2a/(2b-1): a,b \in \mathbb{N} \text{ and } a \geq b\}$ is dense in the rationals $\mathbb{Q}_{>1}$, and hence the reals $\mathbb{R}_{>1}$, as any rational $c/d$ with $c>d$ can be approximated by  $c2^n / (d2^n-1) \in F$ for suficiently large $n$.

@@

\subsection{Renyi divergences on quasi-distributions}
Now we define the classical $\alpha$-Renyi divergence in terms of the Wigner quasi-distribution of any magic state $\rho$ and the Wigner probability distribution of any equilibrium state $\tau$ as
\begin{equation}
	D_\alpha(W_\rho || W_\tau) := \frac{1}{\alpha - 1} \log \sum_{\bmx} W_{\rho}(\bmx)^\alpha W_{\tau}(\bmx)^{1-\alpha},
\end{equation}
where $\alpha$ can take values of the form $2a/(2b+1)$, for non-negative integers $a,b$ to ensure once again that the argument of the logarithm is positive.

@@

We are now in a position to derive general magic state distillation bounds.

@@

	We have the following family of upper bounds on the distillation rate $R := n'/n$,
	\begin{equation}
		R \leq \frac{D_{\alpha}(W_\rho||W_\tau)}{D_{\alpha}(W_{\rho'}||W_{\tau'})} = \frac{F_\alpha(W_\rho) - F_\alpha(W_\tau)}{F_\alpha(W_{\rho'}) - F_\alpha(W_{\tau'})},
	\end{equation}
	where $\alpha = \frac{2a}{2b-1}$ for any positive integers $a,b$ with $a \geq b$ and $\beta F_\alpha(W_\rho) := D_\alpha (W_\rho || W_\tau) - \log \Z$ is the $\alpha$--free energy. 
\end{theorem}
\begin{proof}
	The bound is a direct consequence of the properties of the $\alpha$--Renyi diveregence, proven in~\cref{thm:Da_props}.

Due to the action of the magic protocol channel, we get
\begin{equation}	
	D_\alpha(W_{\rho^{\otimes n}} || W_{\tau^{\otimes n}}) \geq D_\alpha (W_{\rho'^{\otimes m}} || W_{\tau'^{\otimes m}}).\vspace{10pt}
\end{equation}

We can use the additive property to rewrite this as
\begin{equation}
	n D_\alpha(W_\rho || W_\tau) \geq m D_\alpha(W_{\rho'} || W_{\tau'}).
\end{equation}

Since $\rho' \neq \tau'$, we have $D_\alpha(W_{\rho'} || W_{\tau'}) > 0$, which directly leads to the bound
\begin{equation}
	\frac{m}{n} \leq \frac{D_{\alpha}(W_\rho||W_\tau)}{D_{\alpha}(W_{\rho'}||W_{\tau'})}.
\end{equation}
\end{proof}

We now show the following simple equivalence.
\begin{theorem} A state $\rho$ has Wigner negativity if and only if $H_\alpha (W_\rho) < 0$ for some $\alpha >1$.
\end{theorem}
\begin{proof} 
If we have $H_\alpha (W_\rho) <0 $ for $\alpha >1$ then this implies that $\log \sum_{\x} W_\rho (\x)^\alpha > 1$, however it is known that $H_\alpha$ is always non-negative on probability distributions and therefore $W_\rho(\x)$ must be a strict quasi-distribution with negativity.
